---
title: "Pivot Lobsters"
author: "Jasen Mashal"
date: "1/27/2020"
output: html_document
---

Attaching our packages and reading in data

```{r}
library(tidyverse) #install.packages("tidyverse")
library(broom) #install.packages("broom")

library(readxl)
library(here)
library(skimr) #install.packages("skimr")
library(kableExtra) #install.packages("kableExtra")

# read in lobster data
lobsters <- read_excel(here("data", "lobsters.xlsx"),
                       skip = 4)
```

Viewing our data with skimr

```{r}
library(tidyr) #install.packages("tidyr")
skimr::skim(lobsters) #explicitly stated
skim(lobsters) #not explicitly stated
```

# Pivot Tables
In R, Pivot tables are: 'group_by() %>% summarize()'
'n()' function will count observations
```{r}
load(dplyr)
# group by year and site
lobsters %>% 
  group_by(site, year) %>% 
  summarize(count_by_site_year = n())
```

Summarize by multiple variables

```{r}
siteyear_summary <- lobsters %>% 
  group_by(site, year) %>% 
  summarize(count_by_site_year = n(),
            mean_size_mm = mean(size_mm, na.rm = TRUE),
            sd_size_mm= sd(size_mm, na.rm = TRUE))
```

Table formatting with 'kableExtra::kable()'
```{r}
siteyear_summary %>% 
  kable()
```

There are 'r nrow(lobsters)' total lobsters included in this report.

## Activity
1. build from our analyses: calculate the median lobster size for each site and year