---
title: "Creating graphs with ggplot2"
author: "Jasen Mashal"
date: "1/27/2020"
output: html_document
---

# Plots with ggplot2

We are going to make plots in R using the ggplot2 package from the tidyverse.

We will install these packages:

- tidyverse: suite of packages for wrangling & visualization (ggplot)
- readxl: to read in excel files (in many ways)
- here: to help with filepaths


```{r}
# install.packages("tidyverse")
# library(tidyverse)
library(ggplot2)
library(readr) #install.packages("readr")
# install.packages("readxl")
library(readxl)
# install.packages("here")
library(here)
# install.packages("read_csv")
library(read_csv)
# install.packages("read_excel")
```
```{r echo=FALSE}
ci_np <- read_excel(here("data/ci_np.xlsx"))
# also be written:
ca_np <- read_csv(here("data", "ca_np.csv"))
```

### Make my own first graph!

We're going to make a line plot of year (x-variable) vs visitors (y-variable) in ggplot:
```{r}
ggplot(data = ci_np, aes(x = year, y = visitors)) +
  geom_line()
```

#Update the geom_ to geom_point(), geom_col(), geom_area()

Make a base layer for my CI plots:
```{r}
gg_base <- ggplot(data = ci_np, aes(x = year, y = visitors))

gg_base + geom_point()
```

### Now let's start customizing!

- Line / point colors: color
- Change line type: linetype
- Change point style: pch
- Change transparency: alpha

```{r}
gg_base +
  geom_line(
    color = "purple",
    linetype= "dashed"
  )
```

Activity: create a ggplot scatterplot (geom_point) that has triangle symbols (hint: ?pch). of a non-default size (size = #) and that are an updated color.

```{r}
gg_base + 
  geom_point(
    color="dodgerblue",
    pch = 2,
    size = 4
  )    
```


### Update graph aesthetics based on variables

```{r}
gg_base +
    geom_point(
      aes(size  = visitors, color = visitors),
      pch = 10
    )
```

#Add an argument to update the color based on the year value

### Use gg-themes for overall updates

- theme_light()
- theme_minimal()
- theme_bw()

```{r}
gg_base + 
  geom_area() +
  theme_minimal()
```

### Create a multi-series line plot using data for all CA National Parks

```{r}
ggplot(data=ca_np, aes(x = year, y=visitors, group = park_name)) +
  geom_line(aes(color=park_name))

# Try to update the line colors based on the park_name variable
```

```{r}
my_np_graph <- ggplot(data=ca_np, aes(x=year, y=visitors))+
  geom_point(aes(color = park_name), show.legend=FALSE) +
  facet_wrap(~park_name, scales = "free") +
  theme_minimal()
```

Export my graph as a .png
```{r}
ggsave(plot = my_np_graph, here("figures", "np_graph.png"))
```


